---
title: "Reproduce Song & Spaak (2023)"
subtitle: "Multitrophic assembly: a perspective from modern coexistence theory"
author:
  - name: Chuliang Song
    affiliations:
      - name: Princeton University
      - department: Department of Ecology and Evolutionary Biology
    orcid: 0000-0001-7490-8626
    email: clsong.ecology@gmail.com
    url: https://clsong.com
  - name:  Jürg Spaak
    affiliations:
      - Cornell University
      - RPTU Kaiserslautern-Landau
    orcid: 0000-0001-5157-9188
    email: j.w.spaak@gmail.com
    url: https://jwspaak.com/
format: 
  html:
    toc: true
    self-contained: true
    template-partials:
      - title-block.html
    code-fold: true
---

This file aims to reproduce the main results in the paper _Multitrophic assembly: a perspective from modern coexistence theory_ by Song & Spaak.

# Code preparation
Load R pckages
```{r output=F}
# 'tidyverse' is a collection of R packages designed for data science. It includes packages like ggplot2, dplyr, and tidyr.
library(tidyverse)

# 'here' simplifies building file paths. It constructs paths relative to your project directory.
library(here)

# 'geomtextpath' is used for adding text along paths in 'ggplot2' plots, enhancing the visualization of data.
library(geomtextpath)

# 'patchwork' is a package that allows the combination of multiple 'ggplot2' plots into a composite plot.
library(patchwork)

# 'ggthemr' provides themes for 'ggplot2', allowing for the easy and quick beautification of 'ggplot2' plots.
library(ggthemr)

# 'reticulate' is a package that provides an interface between R and Python, allowing users to run Python code within R.
# It allows the execution of Python code in the R console or in R markdown documents.
# Please refer to https://rstudio.github.io/reticulate/index.html for guidelines on using reticulate.
library(reticulate)
```

Load the python files to compute niche and fitness differences
```{r}
source_python(here("numerical_NFD.py"))
source_python(here("LV_multi_functions.py"))
# Define the LV_model function which takes three arguments: N, mu, and A
LV_model <- function(N, mu, A) {
  # The function calculates and returns the result of 'mu' minus the matrix product of 'A' and 'N'
  mu - A %*% N
}

# Define the calculate_NFD function which takes two arguments: mu and A
calculate_NFD <- function(mu, A) {
  # Try to calculate NFD using the NFD_model function (which isn't defined in this code) with LV_model as its argument
  # along with additional arguments defined in a list
  # If the calculation fails, catch the error and return a list with NA values for ND and "F"
  NFD <- tryCatch(
    {
      NFD_model(LV_model, length(mu), args = list(mu, A), from_R = TRUE)
    },
    error = function(e) {
      list(ND = NA, "F" = NA)
    }
  )
  # Return a list with ND and Fitness values from the NFD result
  list(
    ND = NFD$ND,
    Fitness = NFD$"F"
  )
}
```

# Figure 2
Functions to generate the null models
```{r}
generate_interaction_matrix <- function(n1, n2,
                                        alpha,
                                        noise = F,
                                        noise_level = .01,
                                        sign_predator = 'positive',
                                        noise_choice = 'all') {
  n_tot <- n1 + n2

  if (n1 > 0) {
    A11 <- matrix(alpha[1, 1], ncol = n1, nrow = n1)
    A12 <- matrix(alpha[1, 2], ncol = n2, nrow = n1)
    A21 <- matrix(alpha[2, 1], ncol = n1, nrow = n2)
    A22 <- matrix(alpha[2, 2], ncol = n2, nrow = n2)
    diag(A11) <- 1
    diag(A22) <- 1
  }
  if (noise) {
    if(noise_choice == 'all'){
      A11 <- add_gaussian_noise(A11, noise_level)
      A12 <- add_gaussian_noise(A12, noise_level)
      A21 <- add_gaussian_noise(A21, noise_level)
      A22 <- add_gaussian_noise(A22, noise_level)
    } 
    if(noise_choice == 'Spaak'){
      added_noise <- add_gaussian_noise(A12, noise_level) - A12
      A12 <- A12 + added_noise
      A21 <- A21 + 0.1 * t(added_noise)
    }
  }
  
  A <- rbind(cbind(A11, A12), cbind(A21, A22))

  r1 <- rep(1, n1)
  if(sign_predator == 'positive'){
    r2 <- rep(1, n2)
  } else if(sign_predator == 'null'){
    r2 <- rep(-0.01, n2)
  } else if(sign_predator == 'spaak'){
    r2 <- rep(-0.1, n2)
  } else{
    r2 <- rep(-1, n2)
  }
  
  r <- c(r1, r2)

  r_basal <- r1 - A12 %*% solve(A22) %*% r2
  A_basal <- A11 - A12 %*% solve(A22) %*% A21

  r_predator <- r2 - A21 %*% solve(A11) %*% r1
  A_predator <- A22 - A21 %*% solve(A11) %*% A12

  bind_cols(
    as_tibble_col(
      list(A, A_basal, A_predator),
      column_name = "int_mat"
    ),
    as_tibble_col(list(r, r_basal, r_predator),
      column_name = "int_growth"
    )
  ) %>% 
  mutate(matrix_type = c("community", "basal", "predator"))
}


add_gaussian_noise <- function(mat, noise_level){
  mat + matrix(rnorm(ncol(mat) * nrow(mat), mean = 0, sd = noise_level), nrow = nrow(mat))
}

genereate_setup <- function(type, Nrand = 100) {
  if (type == "lower") {
    setup <- expand_grid(
      snum_basal = 2:total_species,
      snum_predator = 5,
      rand_label = 1:Nrand
    )
  }
  if (type == "higher") {
    setup <- expand_grid(
      snum_basal = 5,
      snum_predator = 2:total_species,
      rand_label = 1:10
    )
  }
  if (type == "total") {
    setup <- tibble(
      snum_basal = 2:19,
      snum_predator = 20 - snum_basal
    ) %>%
      expand_grid(
        rand_label = 1:10
      )
  }
  setup
}

# Null model paramerers
# 'total_species' represents the total number of species in a given context or model
total_species <- 10

# 'mean_a' represents the average interaction strength for the alpha matrix
mean_a <- .3

# 'alpha' is a 2x2 matrix that represents interaction strengths between different species in a model
# This matrix is populated by the 'mean_a' value for interactions within and between species, and negative of 'mean_a' for certain interactions
# The 'byrow = T' argument ensures the matrix is filled by rows (left to right, top to bottom)
alpha <- matrix(c(mean_a, mean_a, -mean_a, mean_a), ncol = 2, byrow = T)
```

## Panels A-F
Simulate how changing the percentage of lower-trophic species affects niche and fitness differences.
```{r cache=T}
type = "lower"

set.seed(1010)
# Generate the setup, calculate equilibrium, and determine feasibility
Figure2_lower_raw <- genereate_setup(type, Nrand = 500) %>% # Generate the setup according to the specified 'type'
  mutate(
    # For each combination of snum_basal and snum_predator, generate an interaction matrix
    int_parameters = map2(
      snum_basal, snum_predator,
      ~ generate_interaction_matrix(.x, .y,
        alpha = alpha, noise = T,
        noise_level = 0.1,
        noise_choice = "Spaak",
        sign_predator = "spaak"
      )
    )
  ) %>%
  unnest(int_parameters) %>%
  mutate(equlibirum = map2(int_growth, int_mat, ~ solve(.y, .x))) %>% # Calculate equilibrium for each row
  rowwise() %>%
  mutate(feasibility = sum(equlibirum < 0) == 0) %>% # Determine if the equilibrium values are feasible (all >= 0)
  ungroup()

# Filter for feasible solutions and calculate NFD
Figure2_lower_NDFD <- Figure2_lower_raw %>%
  filter(feasibility) %>% # Only keep rows where feasibility is TRUE
  mutate(
    NFD = map2(int_growth, int_mat, calculate_NFD) # Calculate Niche and Fitness Differences
  ) %>%
  unnest_wider(NFD) # Unnest the NFD list column into separate columns

# Generate data for the plot
Figure2_lower_plot <- Figure2_lower_NDFD %>%
  rowwise() %>%
  filter(length(ND) > 1) %>%
  ungroup() %>%
  select(snum_basal, snum_predator, matrix_type, ND, Fitness, rand_label) %>%
  gather(measure_type, measure, -snum_basal, -snum_predator, -matrix_type, -rand_label) %>%
  mutate(
    # Create a new variable "species" based on matrix_type
    measure = pmap(
      list(snum_basal, snum_predator, measure, matrix_type),
      function(n0, n1, measure, matrix_type) {
        if (matrix_type == "community") {
          res <- measure %>%
            enframe(value = "measure") %>%
            mutate(species = c(rep("Basal", n0), rep("Predator", n1))) %>%
            select(-name)
        } else if (matrix_type == "basal") {
          res <- measure %>%
            enframe(value = "measure") %>%
            mutate(species = c(rep("Basal", n0))) %>%
            select(-name)
        } else if (matrix_type == "predator") {
          res <- measure %>%
            enframe(value = "measure") %>%
            mutate(species = c(rep("Predator", n1))) %>%
            select(-name)
        }
        res
      }
    )
  ) %>%
  unnest(measure) %>%
  distinct() %>%
  mutate(measure_type = ifelse(measure_type == "ND",
    "Niche difference",
    "Fitness difference"
  )) %>%
  filter(snum_basal > 2) %>%
  mutate(
    # Recode matrix_type for better interpretation
    matrix_type = case_when(
      matrix_type == "basal" ~ "Traditional focus",
      matrix_type == "predator" ~ "Alternative focus",
      matrix_type == "community" ~ "Community focus"
    )
  ) %>%
  mutate(
    # Make sure measure_type and matrix_type are factors and have a defined order
    measure_type = ordered(
      measure_type,
      levels = c(
        "Niche difference",
        "Fitness difference"
      )
    ),
    matrix_type = ordered(
      matrix_type,
      levels = c(
        "Traditional focus",
        "Alternative focus",
        "Community focus"
      )
    )
  )
```

Plot the data
```{r message=FALSE}
ggthemr::ggthemr(palette = "fresh", layout = "clean")
Figure2_lower <- Figure2_lower_plot %>%
  group_by(measure_type, matrix_type) %>%
  group_split() %>%
  map(function(xxx) {
    xxx <- xxx %>%
      mutate(species = ifelse(species == "Basal",
        "Lower-trophic",
        "Higher-trophic"
      )) %>% 
      mutate(species = ordered(species, levels=c(
         "Lower-trophic",
        "Higher-trophic"
      )))
    xxx_matrix_type <- unique(xxx$matrix_type)
    xxx_measure_type <- unique(xxx$measure_type) %>%
      as.character()
    ylabel <- ifelse(xxx_matrix_type == "Traditional focus",
      xxx_measure_type, " "
    )
    color_values <- case_when(
      unique(xxx$species) == 2 ~ c("#EFBB24", "#3A8FB7"),
      unique(xxx$species) == "Higher-trophic" ~ c("#3A8FB7"),
      T ~ c("#EFBB24")
    )

    if (type == "higher") {
      xxx <- xxx %>%
        mutate(matrix_type = case_when(
          matrix_type == "Traditional focus" ~ "Alternative focus",
          matrix_type == "Alternative focus" ~ "Traditional focus",
          T ~ "Community focus"
        ))
    }

    xxx_mean <- xxx %>%
      group_by(matrix_type, snum_basal, measure_type, species) %>%
      summarise(
        measure_mean = mean(measure),
        measure_sd = sd(measure),
        .groups = "drop"
      )

    xxx %>%
      ggplot(aes(snum_basal, measure, color = species)) +
      geom_ribbon(
        data = xxx_mean,
        aes(
          x = snum_basal,
          ymin = measure_mean - measure_sd,
          ymax = measure_mean + measure_sd,
          color = species,
          fill = species,
          group = species
        ),
        alpha = .1,
        linetype = 2,
        inherit.aes = F
      ) +
      geomtextpath::geom_textsmooth(aes(label = species),
        straight = T,
        orientation = "x",
        method = "loess"
      ) +
      facet_grid(species ~ matrix_type, scales = "free") +
      scale_color_manual(
        values = color_values
      ) +
      scale_fill_manual(
        values = color_values
      ) +
      labs(
        x = "# of lower trophic",
        y = ylabel
      ) +
      coord_cartesian(
        clip = "off"
      ) +
      scale_y_continuous(breaks = scales::pretty_breaks(n = 4)) +
      theme(
        legend.position = "none",
        strip.text.x = element_text(size = 12)
      )
  }) %>%
  wrap_plots()

Figure2_lower + plot_annotation(tag_levels = "A") &
  theme(strip.text.y = element_blank())
# ggsave("Fig2_lower.pdf", width = 10, height = 6)
```
## Panels G-L
Simulate how changing the percentage of higher-trophic species affects niche and fitness differences.
```{r cache = T}
type = "higher"
set.seed(1010)

Figure2_higher_raw <- genereate_setup(type, Nrand = 500) %>%
  mutate(
    int_parameters = map2(
      snum_basal, snum_predator,
      ~ generate_interaction_matrix(.x, .y,
        alpha = alpha, noise = T,
        noise_level = 0.1,
        noise_choice = "Spaak",
        sign_predator = "spaak"
      )
    )
  ) %>%
  unnest(int_parameters) %>%
  mutate(equlibirum = map2(int_growth, int_mat, ~ solve(.y, .x))) %>%
  rowwise() %>%
  mutate(feasibility = sum(equlibirum < 0) == 0) %>%
  ungroup()

Figure2_higher_NDFD <- Figure2_higher_raw %>%
  filter(feasibility) %>%
  mutate(
    NFD = map2(int_growth, int_mat, calculate_NFD)
  ) %>%
  unnest_wider(NFD)

Figure2_higher_plot <- Figure2_higher_NDFD %>%
  rowwise() %>%
  filter(length(ND) > 1) %>%
  ungroup() %>%
  select(snum_basal, snum_predator, matrix_type, ND, Fitness, rand_label) %>%
  gather(measure_type, measure, -snum_basal, -snum_predator, -matrix_type, -rand_label) %>%
  mutate(
    measure = pmap(
      list(snum_basal, snum_predator, measure, matrix_type),
      function(n0, n1, measure, matrix_type) {
        if (matrix_type == "community") {
          res <- measure %>%
            enframe(value = "measure") %>%
            mutate(species = c(rep("Basal", n0), rep("Predator", n1))) %>%
            select(-name)
        } else if (matrix_type == "basal") {
          res <- measure %>%
            enframe(value = "measure") %>%
            mutate(species = c(rep("Basal", n0))) %>%
            select(-name)
        } else if (matrix_type == "predator") {
          res <- measure %>%
            enframe(value = "measure") %>%
            mutate(species = c(rep("Predator", n1))) %>%
            select(-name)
        }
        res
      }
    )
  ) %>%
  unnest(measure) %>%
  distinct() %>%
  mutate(measure_type = ifelse(measure_type == "ND",
    "Niche difference",
    "Fitness difference"
  )) %>%
  filter(snum_basal > 2) %>%
  mutate(
    matrix_type = case_when(
      matrix_type == "basal" ~ "Traditional focus",
      matrix_type == "predator" ~ "Alternative focus",
      matrix_type == "community" ~ "Community focus"
    )
  ) %>%
  mutate(
    measure_type = ordered(
      measure_type,
      levels = c(
        "Niche difference",
        "Fitness difference"
      )
    )
  ) %>%
  mutate(
    matrix_type = ordered(
      matrix_type,
      levels = c(
        "Traditional focus",
        "Alternative focus",
        "Community focus"
      )
    )
  )
```

Plot the data
```{r message=F}
ggthemr::ggthemr(palette = "fresh", layout = "clean")
Figure2_higher <- Figure2_higher_plot %>%
  group_by(measure_type, matrix_type) %>%
  group_split() %>%
  map(function(xxx) {
    xxx <- xxx %>%
      mutate(species = ifelse(species == "Basal",
        "Lower-trophic",
        "Higher-trophic"
      ))%>% 
      mutate(species = ordered(species, levels=c(
         "Lower-trophic",
        "Higher-trophic"
      )))

    xxx_matrix_type <- unique(xxx$matrix_type)
    xxx_measure_type <- unique(xxx$measure_type) %>%
      as.character()
    ylabel <- ifelse(xxx_matrix_type == "Traditional focus",
      xxx_measure_type, " "
    )
    color_values <- case_when(
      unique(xxx$species) == 2 ~ c("#EFBB24", "#3A8FB7"),
      unique(xxx$species) == "Higher-trophic" ~ c("#3A8FB7"),
      T ~ c("#EFBB24")
    )

    if (type == "higher") {
      xxx <- xxx %>%
        mutate(matrix_type = case_when(
          matrix_type == "Traditional focus" ~ "Alternative focus",
          matrix_type == "Alternative focus" ~ "Traditional focus",
          T ~ "Community focus"
        ))
    }

    xxx_mean <- xxx %>%
      group_by(matrix_type, snum_predator, measure_type, species) %>%
      summarise(
        measure_mean = mean(measure),
        measure_sd = sd(measure),
        .groups = "drop"
      )

    xxx %>%
      ggplot(aes(snum_predator, measure, color = species)) +
      geom_ribbon(
        data = xxx_mean,
        aes(
          x = snum_predator,
          ymin = measure_mean - measure_sd,
          ymax = measure_mean + measure_sd,
          color = species,
          fill = species,
          group = species
        ),
        alpha = .1,
        linetype = 2,
        inherit.aes = F
      ) +
      geomtextpath::geom_textsmooth(aes(label = species),
        straight = T,
        orientation = "x",
        method = "loess"
      ) +
      facet_grid(species ~ matrix_type, scales = "free") +
      scale_color_manual(
        values = color_values
      ) +
      scale_fill_manual(
        values = color_values
      ) +
      labs(
        x = "# of higher trophic",
        y = ylabel
      ) +
      coord_cartesian(
        clip = "off"
      ) +
      scale_y_continuous(breaks = scales::pretty_breaks(n = 4)) +
      theme(
        legend.position = "none",
        strip.text.x = element_text(size = 12)
      )
  })

Figure2_higher[[2]] + Figure2_higher[[1]] + Figure2_higher[[3]] +
  Figure2_higher[[5]] + Figure2_higher[[4]] + Figure2_higher[[6]] +
  plot_annotation(tag_levels = list(LETTERS[7:12])) &
  theme(strip.text.y = element_blank())
# ggsave('Fig2_higher.pdf', width = 10, height = 6)
```

# Figure 3
Simulate how changing proportions of lower- and higher-trophic species affect niche and fitness differences
```{r}
type = "total"
set.seed(1010)
Figure3_raw <- genereate_setup(type, Nrand = 500) %>%
  mutate(
    int_parameters = map2(
      snum_basal, snum_predator,
      ~ generate_interaction_matrix(.x, .y,
        alpha = alpha, noise = T,
        noise_level = 0.1,
        noise_choice = "Spaak",
        sign_predator = "spaak"
      )
    )
  ) %>%
  unnest(int_parameters) %>%
  mutate(equlibirum = map2(int_growth, int_mat, ~ solve(.y, .x))) %>%
  rowwise() %>%
  mutate(feasibility = sum(equlibirum < 0) == 0) %>%
  ungroup()

Figure3_NDFD <- Figure3_raw %>%
  filter(feasibility) %>%
  mutate(
    NFD = map2(int_growth, int_mat, calculate_NFD)
  ) %>%
  unnest_wider(NFD)

Figure3_plot <- Figure3_NDFD %>%
  rowwise() %>%
  filter(length(ND) > 1) %>%
  ungroup() %>%
  select(snum_basal, snum_predator, matrix_type, ND, Fitness, rand_label) %>%
  gather(measure_type, measure, -snum_basal, -snum_predator, -matrix_type, -rand_label) %>%
  mutate(
    measure = pmap(
      list(snum_basal, snum_predator, measure, matrix_type),
      function(n0, n1, measure, matrix_type) {
        if (matrix_type == "community") {
          res <- measure %>%
            enframe(value = "measure") %>%
            mutate(species = c(rep("Basal", n0), rep("Predator", n1))) %>%
            select(-name)
        } else if (matrix_type == "basal") {
          res <- measure %>%
            enframe(value = "measure") %>%
            mutate(species = c(rep("Basal", n0))) %>%
            select(-name)
        } else if (matrix_type == "predator") {
          res <- measure %>%
            enframe(value = "measure") %>%
            mutate(species = c(rep("Predator", n1))) %>%
            select(-name)
        }
        res
      }
    )
  ) %>%
  unnest(measure) %>%
  distinct() %>%
  mutate(measure_type = ifelse(measure_type == "ND",
    "Niche difference",
    "Fitness difference"
  )) %>%
  filter(snum_basal > 2) %>%
  mutate(
    matrix_type = case_when(
      matrix_type == "basal" ~ "Traditional focus",
      matrix_type == "predator" ~ "Alternative focus",
      matrix_type == "community" ~ "Community focus"
    )
  ) %>%
  mutate(
    measure_type = ordered(
      measure_type,
      levels = c(
        "Niche difference",
        "Fitness difference"
      )
    )
  ) %>%
  mutate(
    matrix_type = ordered(
      matrix_type,
      levels = c(
        "Traditional focus",
        "Alternative focus",
        "Community focus"
      )
    )
  )
```


```{r message=FALSE}
ggthemr::ggthemr(palette = 'fresh', layout = 'clean')
Figure3 <- Figure3_plot %>%
  filter(matrix_type == "Community focus") %>%
  mutate(species = ifelse(species == "Basal",
    "Lower\ntrophic",
    "Higher\ntrophic"
  )) %>%
  mutate(species = ordered(species, levels = c(
    "Lower\ntrophic",
    "Higher\ntrophic"
  ))) %>% 
  mutate(perc = snum_basal / (snum_basal + snum_predator)) %>%
  filter(perc <= .75) %>%
  group_split(species, measure_type) %>%
  map(function(xxx) {
    xxx_mean <- xxx %>%
      group_by(matrix_type, perc, measure_type, species) %>%
      summarise(
        measure_mean = mean(measure),
        measure_sd = sd(measure),
        .groups = "drop"
      )
    ylabel <- xxx$measure_type %>% unique()
    xxx_color <- ifelse(unique(xxx$species) == "Higher\ntrophic",
      "#3A8FB7", "#EFBB24"
    )
    xxx %>%
      ggplot(aes(perc, measure,
        color = species
      )) +
      geom_ribbon(
        data = xxx_mean,
        aes(
          x = perc,
          ymin = measure_mean - measure_sd,
          ymax = measure_mean + measure_sd,
          color = species,
          fill = species,
          group = species
        ),
        alpha = .1,
        linetype = 2,
        inherit.aes = F
      ) +
      geomtextpath::geom_textsmooth(aes(label = species),
        straight = T,
        orientation = "x"
      ) +
      facet_grid(species ~ measure_type, scales = "free") +
      labs(
        x = "Percentage of lower trophic"
      ) +
      coord_cartesian(
        clip = "off"
      ) +
      scale_color_manual(
        values = xxx_color
      ) +
      scale_fill_manual(
        values = xxx_color
      ) +
      ylab(ylabel) +
      theme(
        legend.position = "none",
        strip.text.x = element_text(size = 12)
      )
  }) %>%
  wrap_plots()

Figure3 + plot_annotation(tag_levels = "A") &
  theme(strip.text = element_blank())
ggsave('Fig3.pdf', width = 8, height = 6)
```


# Figure 4
Load the source data from [Sauve & Barraquand 2020](https://besjournals.onlinelibrary.wiley.com/doi/10.1111/1365-2656.13227). Export empirical intrinsic growth rates and interaction matrix with their parametrization.
```{r message = F}
PopParamRef <- read_csv("data/Parameterisation/YearPopParam.csv") %>%
  rename(DD = DDData)

IntParamRefI <- read_csv("data/Parameterisation/YearIntParam_TypeI.csv")

ICRef <- read_csv("data/Parameterisation/SpDensBiomass.csv") %>%
  mutate(InitBiomass_gha = BodyMass_g * InitDensity_Nha)

mat_all <- sapply(c("S", "W"), function(x) NULL)
r_all <- sapply(c("S", "W"), function(x) NULL)
for (season in c("S", "W")) {
  eff <- .1 # Set efficiency parameter

  mat_11 <- IntParamRefI %>%
    filter(SeasonDom == season) %>%
    select(LowerTaxon) %>%
    distinct() %>%
    left_join(
      PopParamRef %>%
        select(Taxon, BETA),
      by = c("LowerTaxon" = "Taxon")
    ) %>%
    mutate(
      i = LowerTaxon,
      j = LowerTaxon
    ) %>%
    select(-LowerTaxon) %>%
    rename(alpha = BETA)

  mat_21 <-
    IntParamRefI %>%
    filter(SeasonDom == season) %>%
    left_join(
      PopParamRef %>%
        select(Taxon, BodyMass),
      by = c("UpperTaxon" = "Taxon")
    ) %>%
    mutate(alpha = ifelse(SeasonDom == "S",
      -eff * G * (1 + E) / BodyMass,
      -eff * G * (1 - E) / BodyMass
    )) %>%
    select(LowerTaxon, UpperTaxon, alpha) %>%
    rename(
      j = LowerTaxon,
      i = UpperTaxon
    )

  mat_12 <- mat_21 %>%
    mutate(alpha = -alpha / eff) %>%
    rename(
      i = j,
      j = i
    )

  mat_22 <-
    IntParamRefI %>%
    filter(SeasonDom == season) %>%
    select(UpperTaxon) %>%
    distinct() %>%
    left_join(
      PopParamRef %>%
        select(Taxon, DD),
      by = c("UpperTaxon" = "Taxon")
    ) %>%
    mutate(
      i = UpperTaxon,
      j = UpperTaxon
    ) %>%
    select(-UpperTaxon) %>%
    rename(alpha = DD)


  mat <- bind_rows(
    mat_11,
    mat_12,
    mat_21,
    mat_22
  ) %>%
    reshape2::acast(i ~ j, value.var = "alpha")

  r <-
    PopParamRef %>%
    filter(Taxon %in% rownames(mat)) %>%
    mutate(r = ifelse(TrophicLevel == "Prey", R, -M)) %>%
    select(Taxon, r) %>%
    arrange(
      factor(Taxon, levels = rownames(mat))
    ) %>%
    pull(r, name = Taxon)
  names(r) <- rownames(mat)

  mat_all[[season]] <- mat
  r_all[[season]] <- r

  write.csv(mat, paste0("data/mat_", season, ".csv"))
  write.csv(r, paste0("data/r_", season, ".csv"))
}
```
Then compute the niche and fitness differences for the species
```{python message = F}
import numpy as np
import pandas as pd
import matplotlib.pyplot as plt
import warnings

from scipy.integrate import solve_ivp

import LV_multi_functions as lmf

mu = {"S": pd.read_csv("data/r_S.csv", index_col=0).values[:, 0],
      "W": pd.read_csv("data/r_W.csv", index_col=0).values[:, 0]}
names = {"S": np.array(list(pd.read_csv("data/r_S.csv", index_col=0).index)),
         "W": np.array(list(pd.read_csv("data/r_W.csv", index_col=0).index))}
A = {"S": pd.read_csv("data/mat_S.csv", index_col=0).values,
     "W": pd.read_csv("data/mat_W.csv", index_col=0).values}


for ind in "SW":
    good = (np.isfinite(mu[ind])  # has actual intrinsic growth rate
            & (np.sum(np.isfinite(A[ind]), axis=1) > 0)  # has effects
            & (np.sum(np.isfinite(A[ind]), axis=0) > 0)  # is affected
            & (np.diag(A[ind]) > 0))  # has actual self limitation

    mu[ind] = mu[ind][good]
    A[ind] = A[ind][good][:, good]
    A[ind][np.isnan(A[ind])] = 0
    names[ind] = names[ind][good]


# compute the trophic level
trophic_level = {}
for ind in "SW":
    consum = A[ind] < 0

    # change to relative consumtion
    rowsum = np.sum(consum, axis=1)
    rowsum[rowsum == 0] = 1  # avoid division by 0

    consum = consum/rowsum[:, np.newaxis]

    trophic_level[ind] = np.linalg.solve(np.eye(len(mu[ind])) - consum,
                                         np.ones(len(mu[ind])))


fig, ax = plt.subplots(2, 2, sharex="row", sharey="row",
                       figsize=(9, 9))


A_NFD,ND,FD, mu_NFD, names_NFD, tl = np.empty((6,2), dtype = "object")
for i, ind in enumerate("SW"):
    sol = solve_ivp(lambda t, N: N*(mu[ind] - A[ind].dot(N)),
                    [0, 100], np.full(len(mu[ind]), 0.01))
    ax[0, i].plot(sol.t, sol.y.T)
    ax[0, i].semilogy()
    ax[0, i].set_ylim([1e-5, 1e5])

    ax[1, i].semilogy(trophic_level[ind], sol.y[:, -1], '.')
    ax[1, i].set_ylim([1e-5, 1e5])

    # define where NFD can be computed
    good = np.where(sol.y[:, -1] > 1e-5)[0]
    A_NFD[i] = A[ind][good[:, np.newaxis], good][np.newaxis]
    mu_NFD[i] = mu[ind][good][np.newaxis]
    names_NFD[i] = names[ind][good]
    tl[i] = trophic_level[ind][good]

    computable = [False]

    while (not computable[0]):
        print(len(mu_NFD[i][0]))
        # compute niche and fitness differences
        computable, sub_equi = lmf.find_NFD_computables(A_NFD[i], mu_NFD[i], check = False)
        good = np.where((sub_equi[0] >= 0).all(axis=0))[0]
        A_NFD[i] = A_NFD[i][:, good[:, np.newaxis], good]
        mu_NFD[i] = mu_NFD[i][:, good]
        names_NFD[i] = names_NFD[i][good]
        tl[i] = tl[i][good]
    with warnings.catch_warnings(record = True):
        warnings.simplefilter("ignore")
        ND[i], FD[i], c, NO_ij, FD_ij, r_i = lmf.NFD_LV_multispecies(A_NFD[i],
                                                               sub_equi, r=mu_NFD[i], check = False)
    
    FD[i] = 1-1/(1-FD[i])
    ax[0, i].set_title(ind)
    ax[0,i].set_xlabel("Time")
    ax[1, i].set_xlabel("Trophic level")
    

ax[1, 0].set_ylabel("Density")
ax[0, 0].set_ylabel("Density")

fig, ax = plt.subplots(1,2, sharex = True, sharey = True)

for i in range(2):
    ax[i].plot(ND[i][0, tl[i] < 1.5], FD[i][0, tl[i] < 1.5], 'bo',
                       label = "basal\nspecies")
    ax[i].plot(ND[i][0, tl[i] > 1.5], FD[i][0, tl[i] > 1.5], 'ro',
                       label = "predator\nspecies")
    
    
    ax[i].set_ylim([-5, 1.5])
    ax[i].set_xlabel("Niche differences")

ax[0].set_ylabel("Fitness differences")
ax[0].set_title("Summer")
ax[1].set_title("Winter")
ax[0].legend()
fig.tight_layout()
```
```{r}
empirical <- bind_rows(
  tibble(
    season = "Summer",
    species = py$names_NFD[[1]],
    ND = c(py$ND[[1]]),
    FD = c(py$FD[[1]])
  ),
  tibble(
    season = "Winter",
    species = py$names_NFD[[2]],
    ND = c(py$ND[[2]]),
    FD = c(py$FD[[2]])
  )
) %>%
  left_join(
    r_all %>%
      enframe(name = "season") %>%
      mutate(season = c("Summer", "Winter")) %>%
      mutate(value = map(value, ~ enframe(., name = "species", value = "r"))) %>%
      unnest(value),
    by = join_by(season, species)
  ) %>%
  mutate(trophic = ifelse(r > 0, "Lower trophic", "Higher trophic")) %>%
  mutate(trophic = ordered(trophic,
    levels = c("Lower trophic", "Higher trophic")
  ))
```

Plot the result
```{r warning = F}
ggthemr::ggthemr("fresh", "clean")
empirical %>%
  select(-r) %>%
  group_by(species) %>%
  filter(n() == 2) %>%
  ungroup() %>%
  rename(
    `Niche\ndifference` = ND,
    `Fitness\ndifference` = FD
  ) %>%
  mutate(season = factor(season, level = c("Winter", "Summer"))) %>%
  gather(key, value, -season, -species, -trophic) %>%
  ggplot(aes(
    x = value, y = key,
    fill = season, color = season
  )) +
  geom_boxplot(fill = "white") +
  facet_grid(~trophic, scales = "free_x") +
  scale_color_manual(
    breaks = c("Summer", "Winter"),
    values = c("#1B813E", "#91989F")
  ) +
  labs(
    x = "",
    y = ""
  ) +
  theme(
    legend.title = element_blank(),
    strip.text.x = element_text(size = 14),
    axis.text = element_text(size = 12),
    axis.title = element_blank(),
    panel.spacing = unit(1, "lines")
  )
```

```{r message = F, warning = F}
library(ggside)
ggthemr::ggthemr("pale")
empirical %>%
  select(-r) %>%
  group_by(species) %>%
  filter(n() == 2) %>%
  ungroup() %>%
  rename(
    `Niche difference` = ND,
    `Fitness difference` = FD
  ) %>%
  mutate(season = factor(season, level = c("Summer", "Winter"))) %>%
  ggplot(aes(
    x = `Niche difference`, y = `Fitness difference`,
    color = trophic
  )) +
  geom_point() +
  facet_wrap(~season, scales = "free") +
  scale_color_manual(
    values = c("#EFBB24", "#3A8FB7")
  ) +
  geom_xsideboxplot(aes(y = trophic), orientation = "y") +
  scale_xsidey_discrete(labels = NULL) +
  geom_ysideboxplot(aes(x = trophic), orientation = "x") +
  scale_ysidex_discrete(labels = NULL) +
  theme(
    aspect.ratio = 1,
    legend.position = "bottom",
    legend.title = element_blank(),
    panel.border = element_rect(fill = NA, size = .1),
    strip.text.x = element_text(size = 12)
  )
```
